---
title: "Analysis of Canadian Population"
author: "By Raman Dhanda"
#date: "2023-04-02"
output: 
  html_document: 
    toc: yes
    theme: readable
    fig_width: 10
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{css, echo=FALSE}

body {
/*
rgb(204,193,168)
rgb(247,243,239)
background-color: rgba(246,247,247,0.3);
}
h1{
text-align: center;
}


```

## Preparing Data

### Load Librairies
```{r}
library(tidyverse)      # All important libraries
library(knitr)          # For printing tables
library(stringr)        # Dealing with strings
library(skimr)          # Data summary
library(mice)           # Dealing with missing data
library(scales)         # to for as non scientific notation
# will be making use of IMPUTATION to deal with missing data.
```


### Load Data, Change Data Types, and Rename the Columns
```{r}
can_pop_raw <- read_csv("Data/Population-estimates-quarterly-CA-PT-Q1-1951_Q1-2023.csv", skip=8)
can_pop <- read_csv("Data/Population-estimates-quarterly-CA-PT-Q1-1951_Q1-2023.csv", skip=8,
                    col_types = list('Canada' = col_number(),
                                     'Newfoundland and Labrador' = col_number(),
                                     'Prince Edward Island' = col_number(),
                                     'Nova Scotia' = col_number(),
                                     'New Brunswick' = col_number(),
                                     'Quebec' = col_number(),
                                     'Ontario' = col_number(),
                                     'Manitoba' = col_number(),
                                     'Saskatchewan' = col_number(),
                                     'Alberta' = col_number(),
                                     'British Columbia' = col_number(),
                                     'Yukon' = col_number(),
                                     'Northwest Territories including Nunavut (Terminated) 5' = col_number(),
                                     'Northwest Territories 6' = col_number(),
                                     'Nunavut 6' = col_number()
                    ))
# Rename all Columns
can_pop <- can_pop %>% 
              rename("QY"="Geography",
              "NL"="Newfoundland and Labrador",
              "PE"="Prince Edward Island",
              "NS"="Nova Scotia",
              "NB"="New Brunswick",
              "QC"="Quebec",
              "ON"="Ontario",
              "MB"="Manitoba",
              "SK"="Saskatchewan",
              "AB"="Alberta",
              "BC"="British Columbia",
              "YT"="Yukon",
              "NT-NU"="Northwest Territories including Nunavut (Terminated) 5",
              "NT"="Northwest Territories 6",
              "NU"="Nunavut 6")
#glimpse(can_pop_raw)
#head(can_pop)
#head(can_pop_raw)
glimpse(can_pop)
```


### Create Year and Quarter Columns
```{r}
can_pop <- transform(can_pop, Year = str_sub(QY,-4,-1))
can_pop <- transform(can_pop, Quarter = str_sub(QY,2,2))
can_pop <- transform(can_pop, QTR = str_sub(QY,1,2))
can_pop %>% 
  select(QY, Year, Quarter, QTR) %>% 
  glimpse()
```


### Convert the Data Types of all numeric Columns to integers
```{r}
can_pop <- transform(can_pop, 
                     Canada = as.integer(Canada),
                     NL = as.integer(NL),
                     PE = as.integer(PE),
                     NS = as.integer(NS),
                     NB = as.integer(NB),
                     QC = as.integer(QC),
                     ON = as.integer(ON),
                     MB = as.integer(MB),
                     SK = as.integer(SK),
                     AB = as.integer(AB),
                     BC = as.integer(BC),
                     YT = as.integer(YT),
                     NT.NU = as.integer(NT.NU),
                     NT = as.integer(NT),
                     NU = as.integer(NU),
                     Year = as.integer(Year),
                     Quarter = as.integer(Quarter)
                         )
glimpse(can_pop)
```



### Dealing with missing Data
#### Visual representation of missing data
```{r}
#skim(can_pop)
md.pattern(can_pop)
```

All columns have 2 missing data points expect 3 territories: NT-NU, NT and NU. Let's plot these 3 columns along with YT to see what is happening.


#### Plot Columns with missing data
```{r}
ggplot()+
  geom_line(data=can_pop, mapping=aes(Year, YT), colour="red")+
  geom_line(data=can_pop, mapping=aes(Year, NT.NU), colour="blue")+  #using NT-NU does not work
  geom_line(data=can_pop, mapping=aes(Year, NT), colour="black")+
  geom_line(data=can_pop, mapping=aes(Year, NU), colour="green")+
  geom_smooth()
  
```

#### Identify where the split accurs.
```{r}
can_pop %>% 
  select(Year, QTR, Quarter, YT, NT.NU, NT, NU) %>% 
  drop_na(NT,NU) %>% 
  head()
```

This shows that in 1991 Q4, NT-NU was split into NT and NU.

Before deciding what could be done with this, lets compare the population of these 4 territories.

Lets combine the 3 Territories into one group of Territories
```{r}
can_pop_group <- can_pop %>% 
                  select(everything()) %>% 
                  mutate(TR_GR = if_else((Year>1991) | 
                                          ((Year==1991) & (Quarter==4)),
                                            (YT + NT + NU),
                                            (YT + NT.NU)))
can_pop_group %>% 
  select(QY,Year, Quarter, QTR, TR_GR) %>% 
  head()
```

### Reorder the columns
```{r}
can_pop_group <- can_pop_group %>% 
  select(QY, Year, Quarter, QTR, Canada, NL, PE, NS, NB, QC, ON, MB, SK, AB, BC, TR_GR)
head(can_pop_group)

```

Pivot the Table to long form. This will make it easy to work with.
```{r}
can_pop_group %>% 
  pivot_longer(cols = 5:16,
               names_to = "PT",
               values_to = "Pop") %>%
  filter((Year==1955 | Year == 1990 | Year == 2023) & Quarter==1) %>% 
  select(Year, PT, Pop) %>%
  filter(PT!='Canada') %>% 
        ggplot(aes(x=reorder(PT,-Pop), y=Pop, fill=factor(Year)))+ # reorder PT per Population values and use factor(Year) to have side by side PT values for different years
        geom_bar(stat="identity",position=position_dodge())+    # to not plot the count but the `value
        scale_fill_brewer(palette="Paired")+
        scale_y_continuous(labels = comma)+      # to show values as ##,###,###
        geom_text(aes(label=Pop), angle=0, position=position_dodge(0.7), vjust="right", hjust="center", size=2)
```







```{r, echo=FALSE, include=FALSE}
glimpse(can_pop)
```


### Nice Table
```{r, echo=FALSE, include=FALSE}
can_pop %>% 
  select(Year, Canada, ON, BC) %>% 
  head() %>% 
  kable()
```


####
```{r, include=FALSE}
ggplot(data = can_pop,
       aes(x=Year,
           y=Canada))+
  geom_point(size=2)+
  geom_line(colour="red", size=1)
  geom_line()
#head(can_pop,10)
#glimpse(can_pop)
#skim(can_pop)
#skim(can_pop)
```




